### 2020-04-15 samplepaper.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\input{preamble}
%\usepackage{times}
\usepackage[moderate]{savetrees}
\usepackage{multicol}

\usepackage{graphicx}

% Several figures with different labels
\usepackage{float}
\restylefloat*{figure}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Approaching Optimal Duplicate Detection in a Sliding Window}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Rémi Géraud--Stewart\inst{1} \and
Marius Lombard-Platet\inst{1,2} \and
David Naccache\inst{1}}
%
%\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Département d'informatique de l'ENS, ENS, CNRS, PSL Research University, Paris, France \and
Be-ys Research Lab, Clermont-Ferrand, France\\
\email{\{first\}.\{last\}@ens.fr}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Duplicate detection is the problem of identifying whether a given item has previously appeared in a (possibly infinite) stream of data, when only a limited amount of memory is available.

%Building on the Bloom filter \cite{bloom_spacetime_1970}, an early data structure amenable to duplicate detection, several generalizations and variants were proposed that reduce error rates while maintaining constant memory.

Unfortunately the infinite stream setting is ill-posed, and error rates of duplicate detection filters turn out to be heavily constrained.%: as a consequence they appear to provide no advantage, asymptotically, over a biased coin toss \cite{10.1145/3297280.3297335}.

%Both as a better-behaved setting and as a construction technique, a
%\emph{windowed} variant of the duplicate detection problem was 
%introduced in \cite{She08,Yoo10} which consists in finding duplicate 
%entries amongst the last $w$ elements of the stream for a fixed integer $w$.

In this paper we formalize the sliding window setting introduced by  \cite{She08,Yoo10}, and show that a perfect (zero error) solution can be used up to a maximal window size $w_\text{max}$. Above this threshold we show that some existing duplicate detection filters (designed for the \emph{non-windowed} setting) perform better that those targeting the windowed problem. Finally, we introduce a \enquote{queuing construction} that improves on the performance of some duplicate detection filters in the windowed setting.

We also analyse the security of our filters in an adversarial setting. 
\keywords{TODO}
\end{abstract}
%
%
%
\input{article_content}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
\bibliography{bibli_short}
%\bibliography{qhtv2}

\appendix
\input{appendix}

\end{document}