### 2020-08-19 slides/graphs/bench_short.tex
slides

\begin{tikzpicture}
\begin{semilogxaxis}[%legend pos = outer north east, 
	legend pos= north east,%
	xlabel=$w$, ylabel=$\FPR^w+\FNR^w (\times 100)$, xmin=40,ymin=0,mlineplot]

\addplot[color=blue, mark=o] table[x=w,y=Error] {graphs/shf.dat};

\addplot [draw=none,pattern=north west lines,pattern color=black,domain=10:1843,stack plots=y] {100} \closedcycle;
\addlegendentry{SHF}

%\addplot[color=red, mark=x] table[x=a,y=b] {graphs/CSHF.dat};
%\addlegendentry{CSHF}

\end{semilogxaxis}
\end{tikzpicture}