### 2020-08-19 slides/main.tex

\documentclass[aspectratio=169]{beamer}

\usepackage[utf8]{inputenc}
%\usepackage{kotex}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usepackage{xspace}

%\usepackage[sfdefault]{FiraSans}
%\usepackage{FiraMono}

% Plots
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{fillbetween}
\tikzset{mynode/.style={circle,draw,font=\small,minimum width=1cm,font=\sffamily}}
\pgfdeclarepatternformonly{north east lines wide}%
{\pgfqpoint{-1pt}{-1pt}}%
{\pgfqpoint{10pt}{10pt}}%
{\pgfqpoint{6pt}{6pt}}%
{
	\pgfsetlinewidth{0.1pt}
	\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
	\pgfpathlineto{\pgfqpoint{6.1pt}{6.1pt}}
	\pgfusepath{stroke}
}
\newcommand*{\FPR}{\operatorname{FPR}}
\newcommand*{\FNR}{\operatorname{FNR}}
\newcommand*{\FN}{\operatorname{FN}}
\newcommand*{\FP}{\operatorname{FP}}
\newcommand*{\EP}{\operatorname{EP}}

\newcommand*{\DUPLICATE}{\textsf{DUPLICATE}\xspace}
\newcommand*{\UNSEEN}{\textsf{UNSEEN}\xspace}

\usepackage{graphicx}

\usetheme{metropolis}           % Use metropolis theme

\titlegraphic{%
	\hfill\includegraphics[width=.3\textwidth]{logo-institute.png}\hfill
}

%\titlegraphic{\hfill \includegraphics[width=.2\textwidth]{logo-institute.png}\hfill}
\title{Approaching Optimal Duplicate Detection in a Sliding Window}
\date{August 30th, 2020}
\author{Rémi Géraud-Stewart, \textbf{Marius Lombard-Platet}, David Naccache}
\institute{Information security group, DIENS, École normale supérieure, PSL Research University (Paris, France)}

\begin{document}
\begin{frame}
	\maketitle
\end{frame}
\section{Introduction and motivation}

\begin{frame}{Finding duplicates}
	Let us consider a stream $E = (e_1, \dotsc, e_n, \dotsc)$ of elements, and a machine with a limited amount of memory $\mathcal M$.

	\begin{alertblock}{Question}
		How to detect if an element $e$ has already been seen in the past?
	\end{alertblock}%\pause

%	\begin{alertblock}{Second question}
%		Why do we care?
%	\end{alertblock}
%
%	\begin{exampleblock}{}
%			\begin{itemize}
%			\item Cache for CPU or webservers
%			\item Fraud detection (identical payments)
%			\item Cryptography (nonce reuse)
%			\item Security (check if URL is in a denylist)
%			\item ...
%		\end{itemize}
%	\end{exampleblock}

\end{frame}

\begin{frame}{Example}
This question leads to a second question: \pause

Why do we care?
\pause

There are several applications of detecting duplicates.
\begin{itemize}
	\item Cache for CPU or webservers
	\item Fraud detection (identical payments)
	\item Cryptography (nonce reuse)
	\item Security (check if URL is in a denylist)
	\item ...
\end{itemize}
\end{frame}

\begin{frame}{What is a duplicate filter?}
Quickly said, a duplicate filter is:
\begin{itemize}
	\item An algorithm operating with a finite amount of memory \pause
	\item When presented to a new element, outputs $\DUPLICATE$ or $\UNSEEN$ \pause
	\item Tries to keep memory of the previous elements
\end{itemize}
\end{frame}

\begin{frame}{Sub-problems of duplicate detection}
	In the literature, we can find several variants of the duplicate detection problem (DDP). We can match an element against:\pause
	\begin{itemize}
		\item a fixed, non-evolutive set (fixed duplicate detection, FDDP) \pause
		\item an evolutive set (streaming duplicate detection problem, SDDP) \pause comprising of:
		\begin{itemize}
			\item {\color<6->{red}the last $w$ elements of the stream ($w$DDP)} \pause
			\item all previous elements of the stream ($\infty$DDP)
		\end{itemize} 
	\end{itemize}
\pause
\end{frame}

\bgroup
\let\oldfootnoterule\footnoterule
\def\footnoterule{\only<2->\oldfootnoterule}
\begin{frame}{Prior work}
	Perfect solutions for $\infty$DDP and FDDP on an alphabet of size $\Gamma$ requires $\Gamma$ bits of memory\pause

	One can significantly reduce the amount of memory while keeping the errors under a certain value\footnote<2->{Burton H. Bloom. 1979. \textit{Space/Time Trade-offs in Hash Coding with Allowable Errors}} \pause

	Asymptotically, no filter can perform better than random on the $\infty$DDP\footnote<3->{Rémi Géraud, Marius Lombard-Platet, and David Naccache. 2019. \textit{Quotient hash tables: efficiently detecting duplicates in streaming data}}

\end{frame}

\begin{frame}{Contributions}
	We introduce a lower bound on the error rate for $\infty$DDP \pause

	We introduce two new structures for the $w$DDP:
	\begin{itemize}
		\item One for small values $w$
		\item One for adapting $\infty$DDP algorithms to the $w$DDP \pause
	\end{itemize}

	For these structures, we analyse their error rate and their resistance to adversarial attacks.
\end{frame}
\egroup

\section{Limitations of the $\infty$DDP}

\begin{frame}{Lower bound on $\infty$DDP error rate}
	Let be a filter of $M$ bits of memory on an alphabet of size $\Gamma$

	For $\FP_n, \FN_n$ the probability of a false positive/negative after $n$ insertions, and $\EP_n = \FN_n + \FP_n$ the error probability,\pause

	$$
	EP_n \geq 1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n} 
	$$

	We observe that asymptotically, $\EP_\infty \gtrapprox 1 - \frac M\Gamma$

\end{frame}
\begin{frame}{Experimental saturation}
	Assuming a large alphabet of size $U$, a filter quickly saturates, to the point of answering randomly (FPR + FNR = 1)

	\begin{figure}
		\centering
		\scalebox{.7}{\begin{tikzpicture}
		\input{graphs/er_n_artificial}
		\end{tikzpicture}}
		\caption{Error rate (times 100) of DDFs of 1Mb as a function of stream length. Hatched area represents over-optimal (impossible) values.}
	\end{figure}
\end{frame}

\section{New Structures for the $w$DDP}

\begin{frame}{How to detect duplicates?}
	\textbf{Optimal detection}: $w\log_2 \Gamma$ bits.\pause

	\textbf{Initial idea}: all $w$ elements in a queue \pause
	\begin{itemize}
		\item optimal size, no error
		\item complexity increases with $w$
	\end{itemize}
\pause
	\textbf{Second idea}: add a counter for each element \pause
	\begin{itemize}
		\item O(1) complexity while keeping a constant size
		\item no errors
	\end{itemize}\pause

	Problem solved?
\end{frame}
\begin{frame}{Optimal detection}
	Let $\mathcal M$ the available memory, $w$ the size of the sliding window and $\Gamma$ the size of the alphabet, \pause

	\begin{itemize}
		\item If $M \geq w(\log_2(w) + 2\log_2(\Gamma))$ the wDDP can be perfectly solved in constant time. \pause

		\item If $M \geq 5w\log_2(w)$, the wDDP can be solved with an FPR of $\approx \frac 1w$, and FNR of $0$ and constant time.
	\end{itemize}

\end{frame}

\begin{frame}{Detection with smaller-than-optimal memory}
	As we just saw, detection is optimally solved when the memory is large enough. \pause

	So let's see what happens with smaller memory (or bigger $w$).
\end{frame}

\begin{frame}{A first try: SHF}
	We take the previous optimal solution: a queue and a counter \pause

	Rather than each element $e$, we store a \emph{hash} $h(e)$, of size $\lfloor \frac M{2w} - \frac 12 \log_2(w)\rfloor$. \pause 
	This value is chosen so that we can store a hash of each element of the sliding window.

	That's the SHF. \pause

	\begin{itemize}
		\item $\FN = 0$
		\item$\FP = 1 - \left(1 - \sqrt{w2^{-M/w}}\right)^w$
	\end{itemize}
\end{frame}

\begin{frame}{SHF saturation}

	\begin{columns}
		\begin{column}{0.5\textwidth}
	SHF saturates quickly after reaching a turning point.

	\vspace{6em}
	\textbf{Conclusion}: SHF can be used for optimal cases and a bit more.
		\end{column}
		\begin{column}{0.5\textwidth}  %%<--- here
			\begin{figure}[t]
			\centering
			\scalebox{.7}{
				\input{graphs/bench_short.tex}
			}
			\caption{Error rate (times 100) of SHF on various $w$ sizes. Hatched area represents cases where optimal solutions exist.}
		\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Another idea: Queuing Filter}
	The idea is to adopt $\infty$DDP filters into the $w$DDP context. For this, we create a queue of subfilters $\mathcal F_i$. \pause

	\begin{figure}
		\centering
		\scalebox{.8}{
			\input{graphs/queuef.tex}
		}
	\end{figure}
\onslide<+->{We chose the parameters $L, c$ so that $Lc \approx w$.}
\end{frame}

\begin{frame}{Queueing Filter error probabilities}
	\begin{columns}
		\begin{column}{0.5\textwidth}
		FP and FN probabilities have been derived for a queueing filter using any kind of subfilter, but are a bit complicated to use. \pause

	\vspace{1em}

	Let's rather look at experiments.
		\end{column}
		\begin{column}{0.5\textwidth}  %%<--- here
			\begin{figure}[t]
			\centering
			\scalebox{.7}{
				\input{graphs/queue_bench.tex}
			}
			\caption{QHT-queuing filter and a QHT filter on various sliding window sizes. Hatched area represents values for which optimal solution exists.}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\section{Queueing Filter Security}
\begin{frame}{Filters in adversarial environment}
	An adversary chooses the stream sent to the filter \pause

	She does not have access to the memory state, only to the filter's output ($\DUPLICATE$ or $\UNSEEN$) \pause

	She wins the $(p, n)$-adversarial game if after a training period not longer than $n$, she can force a false positive (or a false negative) with probability $\geq p$
\end{frame}

\begin{frame}{FP security of queueing filter}
	Let $\mathcal Q$ be a queueing filter consisting of $L$ subfilters $\mathcal F_i$, each filled with $c$ items.

	If $\mathcal F$ is $(p, c)$-resistant to adversarial FP attacks and $cL \leq w$, then $\mathcal Q$ is $(1 - (1 - p)^L, w)$-resistant to FP attacks.
\end{frame}

\begin{frame}{FN security of queueing filter}
	Let $\mathcal Q$ be a queueing filter consisting of $L$ subfilters $\mathcal F_i$, each filled with $c$ items, and assume $w \leq (L-1)c$.

	Let $q = \displaystyle\min_{i \leq c} \FN_{\mathcal F, i}$, then $\mathcal Q$ is $(\min(1-q, p)^{L-1}p, w)$-resistant to adversarial FN attacks.
\end{frame}

\section{Conclusion}

\begin{frame}{Summing it up}
	\begin{itemize}
		\item For most cases, $\infty$DDP is an ill-posed problem and existing solutions are close to optimality \pause
		\item $w$DDP can be optimally solved if the memory is big enough \pause
		\item For w close (or inside) optimality, a excellent solution is the SHF \pause
		\item For $w$ slightly bigger, queueing filters are more interesting \pause
		\item For $w$ bigger, then existing $\infty$DDP solutions are the best solution
	\end{itemize}
\end{frame}

\begin{frame}[standout]{Thank you for your attention}
Any question?

\vspace{2em}

\footnotesize \texttt{remi.geraud@ens.fr\\marius.lombard-platet@ens.fr\\david.naccache@ens.fr}
\end{frame}

\end{document}